---
title: "Creating Dumbbell Charts with ggcharts"
author: "Thomas Neitmann"
output: xaringan::moon_reader
---
<style>
.remark-slide-number {
  display: none;
}
</style>

```{r setup, echo=FALSE}
knitr::opts_chunk$set(dpi = 300, fig.ext = "svg", dev = "svglite")
```

## Dumbbell Charts with ggcharts

```{r dumbbell_chart, echo=FALSE, message=FALSE, fig.height=4.5}
library(ggcharts)
library(dplyr)
library(gapminder)
data(gapminder)

gapminder %>%
  dplyr::mutate(pop = pop / 1e6) %>%
  dplyr::filter(year %in% c(1952, 2007)) %>%
  tidyr::pivot_wider(
    id_cols = country,
    names_from = year,
    values_from = pop,
    names_prefix = "pop_"
  ) %>%
  dumbbell_chart(country, pop_1952, pop_2007, limit = 10,
                 point_colors = c("lightgray", "black"),
                 legend_labels = c("1952", "2007")) +
  scale_y_continuous(expand = expand_scale(mult = .025)) +
  theme(
    text = element_text(size = 14),
    plot.caption = element_text(family = "mono")
  ) +
  labs(
    title = "The World is Growing",
    subtitle = "Top 10 Most Populated Countries in 2007",
    caption = "Chart created using #ggcharts"
  ) +
  xlab(NULL) +
  ylab("Population (Millions)")
```

---

## Data Preparation

For this demo I will use the `gapminder` dataset which contains countries' population counts from 1952 to 2017. This dataset is in long format. In order for `dumbbell_chart()` to work properly the data has to be in *wide* format.

```{r transpose_data}
gapminder_wide <- gapminder::gapminder %>%
  dplyr::mutate(pop = pop / 1e6) %>%
  dplyr::filter(year %in% c(1952, 2007)) %>%
  tidyr::pivot_wider(
    id_cols = country,
    names_from = year,
    values_from = pop,
    names_prefix = "pop_"
  )
```

---

## Data Preparation (cont'd)

```{r preview_data}
head(gapminder_wide, 10)
```

---

## A Simple Chart

```{r simple_chart, fig.height=3.75}
library(ggcharts)
dumbbell_chart(gapminder_wide, country, pop_1952, pop_2007,
               limit = 10)
```

---

## Customize

```{r customize1, fig.height=3.5}
chart <- dumbbell_chart(gapminder_wide, country, pop_1952, pop_2007,
                        limit = 10, point_size = 5,
                        point_colors = c("lightgray", "black"),
                        legend_labels = c("1952", "2007"))
chart
```

---

## Customize (cont'd)

```{r customize2}
chart <- chart +
  scale_y_continuous(expand = expand_scale(mult = .025)) +
  theme(
    text = element_text(size = 14),
    plot.caption = element_text(family = "mono")
  ) +
  labs(
    title = "The World is Growing",
    subtitle = "Top 10 Most Populated Countries in 2007",
    caption = "Chart created using #ggcharts"
  ) +
  xlab(NULL) +
  ylab("Population (Millions)")
```

---

## Final Result

```{r final, fig.height=4.25}
chart
```
